---
title: "Representativeness of extended survey"
output: html_notebook
---

```{r setup}
library(tidyverse)
library(Hmisc)

theme_set(theme_minimal())

# Fix default ggsave background
ggsave <- function(..., bg = 'white') ggplot2::ggsave(..., bg = bg)

figdir <- "figures/manuscript/supplementary"
```

Aim: to understand how participants of the extended survey are representative of:
+ participants of the primary survey
+ residents in surveyed households (household survey)
+ residents of survey clusters overall (worldpop)

```{r load_data}

# Raw survey data
datadir <- "C:/Users/phpuenig/Dropbox/SCALE_Hit-TB_modelling-resources/ModellingData/Pre_PS_Data"
ind <- readRDS(file.path(datadir, "individual_clean.rds"))
hh <- readRDS(file.path(datadir, "household_clean.rds"))

# Analysis data
source(here::here("code/load_data.R"), echo=TRUE)
dat_nona <- readRDS(here::here("data","analysis_data_primary.rds"))

clust <- readRDS(here::here("data","dat_scale.rds")) %>% 
  sf::st_drop_geometry() 

# Unadjusted fits
fit_base <- readRDS(here::here("output","fit_base.rds"))
fit_list_pov <- readRDS(here::here("output","fit_list_pov.rds"))

priors <- c(prior(normal(-2, 2), class = Intercept),
            prior(normal(0,10), class = b),
            prior(cauchy(0,1), class = sd, group = "clustid"))

```

# Summarise each sample by age and sex

## Household survey

Reported population across sampled households.

```{r household_pop}

hh %>% 
  dplyr::select(hh_id, h07hh_number:h09hh_minor_num, starts_with("h10"), starts_with("hh11")) -> hh_

# Total household members
sum(hh$h08hh_adults_num) # 20555
n_distinct(hh$hh_id) # 7175

hh_ %>% 
  dplyr::select(!starts_with("hh11")) %>% 
  pivot_longer(starts_with("h10")) %>% 
  rename(age = value) %>% 
  mutate(name = gsub("h10membr_age","",name)) -> hh_age_long

hh_ %>% 
  dplyr::select(!starts_with("h10")) %>% 
  pivot_longer(starts_with("hh11")) %>% 
  rename(sex = value) %>% 
  mutate(name = gsub("hh11membr_sex","",name)) -> hh_sex_long

hh_long <- hh_age_long %>% 
  full_join(hh_sex_long) %>% 
  filter(!(is.na(age & is.na(sex)))) 

# 88 appears to be a missing code
hist(hh_long$age)

hh_long <- hh_long %>% 
  mutate(age = ifelse(age < 0, NA, age),
         age = na_if(age, 88))

hist(hh_long$age)
summary(hh_long)
# No missing sex

hh_long <- hh_long %>% 
  group_by(hh_id) %>%
  mutate(hh_tot = sum(!is.na(sex))) %>%
  filter(!is.na(sex)) %>%
  mutate(sex = factor(sex, levels = 1:2, labels = c("Male","Female")),
         hh_tot_match = (h07hh_number == hh_tot),
         agegp = factor(cut(age, 
                            breaks = c(17,24,34,44,54,120)),
                        levels = c("(17,24]","(24,34]","(34,44]","(44,54]","(54,120]"))) 
  
hh_dist <- hh_long %>% 
  # exclude minors
  filter(!is.na(agegp)) %>% 
  group_by(sex, agegp) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(survey = "Household")

```

## Individual survey

Calculate totals by age and sex for primary and extended survey samples

```{r ind_survey}

# First redefine age groups
ind <- mutate(ind,
              agegp = factor(cut(s09age, 
                                 breaks = c(17,24,34,44,54,120)),
                        levels = c("(17,24]","(24,34]","(34,44]","(44,54]","(54,120]")))

# Primary
ind %>% 
  group_by(sex, agegp) %>% 
  count() %>% 
  # Calculate distribution across age groups for male/female
  # group_by(sex) %>% 
  # mutate(p = n/n()) %>% 
  ungroup() %>%
  mutate(survey = "Primary") -> ind_dist
ind_dist

# Extended
ind %>% 
  filter(!is.na(s50tb_medtest)) %>% 
  group_by(sex, agegp) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(survey = "Extended") -> ext_dist
ext_dist

```

## SCALE survey clusters (Mcewen's population estimates)

```{r scale_clusters}

clust %>% 
  summarise(across(male_adults:female_adults, sum),
            survey = "Cluster total") %>% 
  pivot_longer(-survey, names_to = "sex", values_to = "n") %>% 
  mutate(sex = stringr::str_to_sentence(gsub("_adults","",sex))) -> clust_dist

```

# Totals

```{r calc_totals}

sum(clust$total)
sum(hh_dist$n)
sum(ind_dist$n)
sum(ext_dist$n)

```

# Compare by sex

What proportion female across surveyed clusters, sampled households, and survey participants?

```{r  compare_sex}

clust_dist %>%  
  bind_rows(hh_dist) %>% 
  bind_rows(ind_dist) %>% 
  bind_rows(ext_dist) %>% 
  group_by(survey, sex) %>% 
  summarise(n = sum(n)) %>% 
  group_by(survey) %>% 
  mutate(N = sum(n)) %>% 
  ungroup() %>% 
  mutate(p = n/N,
         low = binconf(n, N, method = "exact")[,2],
         hi = binconf(n, N, method = "exact")[,3]) %>% 
  mutate(across(p:hi, function(x)x*100),
         survey = factor(survey, 
                         levels = c("Cluster total","Household","Primary","Extended"), 
                         labels = c("Survey area\ntotal",
                                    "Household\nsurvey",
                                    "Primary\nsurvey",
                                    "Extended\nsurvey"))) -> bysex

bysex %>% 
  filter(sex == "Female") %>% 
  ggplot(aes(x = survey, y = p, ymin = low, ymax = hi)) +
  geom_errorbar( width = 0.2)+
  geom_point() + 
  labs(x = "", y = "% Female", col = "") +
  theme(text = element_text(size = 14)) -> prop_female

prop_female
# ggsave(here::here(figdir, "compare_prop_female.png"), height = 4, width = 5)

```

# Compare by age

```{r compare_primary_extended1}

ind_dist %>% 
  bind_rows(ext_dist) %>%
  filter(!is.na(agegp)) %>% 
  group_by(survey, agegp) %>% 
  summarise(n = sum(n)) %>% 
  group_by(survey) %>% 
  mutate(p = n/sum(n)) %>%
  mutate(p = ifelse(survey == "Primary", p*(-1), p*1)) %>%
  ggplot(aes(x = agegp, y = p, fill = survey)) + 
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_y_continuous(limits = c(-0.4, 0.4)) +
  labs(fill = "", y = "Proportion", x = "Age group") 

hh_dist %>% 
  bind_rows(ext_dist) %>%
  filter(!is.na(agegp)) %>% 
  group_by(survey, agegp) %>% 
  summarise(n = sum(n)) %>% 
  group_by(survey) %>% 
  mutate(p = n/sum(n)) %>%
  mutate(p = ifelse(survey == "Household", p*(-1), p*1)) %>%
  ggplot(aes(x = agegp, y = p, fill = survey)) + 
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_y_continuous(limits = c(-0.4, 0.4)) +
  labs(fill = "", y = "Proportion", x = "Age group") 

# ggsave(here::here(figdir, "hh_extended_age_pyramid.png"), height = 5, width = 5)

```
Proportions falling into each age group are roughly similar between the household and sub-sampled extended survey, but survey has a higher proportion of younger participants. 

# Compare by age and sex

```{r hh_byagesex}

compare <- ind_dist %>%
  bind_rows(ext_dist) %>% 
  bind_rows(hh_dist) %>% 
  filter(!is.na(agegp)) %>% 
  mutate(survey = factor(survey, levels = c("Household","Primary","Extended")))
  
compare %>% 
  filter(survey != "Primary") %>% 
  mutate(n = ifelse(sex == "Male", n*(-1), n*1)) %>% 
  ggplot(aes(x = agegp, y = n, fill = sex)) + 
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(fill = "", y = "Frequency", x = "Age group") +
  facet_wrap(~survey, scales = "free_x") +
  theme(text = element_text(size = 14))

compare %>% 
  filter(survey != "Primary") %>% 
  group_by(survey) %>% 
  mutate(p = n/sum(n)) %>% #View
  ggplot(aes(x = agegp, y = p, col = survey, group = survey)) + 
  geom_line() +
  geom_point() +
  facet_wrap(~sex) +
  theme(legend.position = c(0.9, 0.8)) +
  labs(fill = "", y = "Proportion", x = "Age group", col = "") +
  theme(text = element_text(size = 14)) -> compare_age_sex


prop_female + compare_age_sex + plot_annotation(tag_levels = "A") +
  plot_layout(widths = c(2,3))

ggsave(here::here(figdir, "figs1.png"), height = 4, width = 12)

```


```{r compare_hh_extended}

compare %>%
  filter(!is.na(agegp)) %>% 
  group_by(survey, agegp) %>% 
  summarise(Male = sum(n[sex == "Male"]),
            Female = sum(n[sex == "Female"])) %>% 
  pivot_longer(-c(survey, agegp)) %>% 
  mutate(value = ifelse(name == "Male", value*(-1), value*1)) %>%
  ggplot(aes(x = agegp, y = value, fill = name)) + 
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(fill = "", y = "Frequency", x = "Age group") +
  facet_wrap(~survey, scales = "free_x")

# ggsave(here::here(figdir, "sample_age_sex_pyramid.png"), height = 5, width = 6)

```

# Post-stratification

Relative to the household survey population, by age group and sex.

Refit primary model to be multilevel, with random intercepts on each of the post-strat characteristics.

```{r refit_mlm}

fit_base_ml <- brm(test01 ~ factor(sex) + factor(sex)*factor(agegp) + 
                     (1 | agegp) + (1 | hiv) + (1 | clustid), 
                   data = dat_nona,
                   prior = priors,
                   family = "bernoulli")

```

```{r ps}

hh_dist %>% 
  mutate(prop = n/sum(n),
         freq = 72) %>% 
  tidyr::uncount(freq) %>% 
  mutate(clustid = rep(1:72, 10)) -> ps_data

ps_epred <- posterior_epred(fit_base, newdata = ps_data)

```


