---
title: "Baseline model development"
output: html_notebook
---

```{r setup}

set.seed(0000)

library(tidyverse)
library(gtools)
library(ggmap)
library(sf)
library(patchwork)
library(brms)
library(spdep)
library(tidybayes)
library(sjPlot)
library(bayesplot)

# devtools::install_github("m-clark/visibly")
# library(visibly)

theme_set(theme_minimal())

figdir <- "figures/fit"

# Fix default ggsave background
ggsave <- function(..., bg = 'white') ggplot2::ggsave(..., bg = bg)

#------------------------------------------------------------------------------#
# Run data loading script

source(here::here("code/load_data.R"), echo=TRUE)

```

# Setup 

## Study area

```{r area_map}

# Map clusters and clinic areas

clinicarea$type = "Clinic"
clust$type = "Cluster"

clust %>% 
  bind_rows(clinicarea) %>% 
  mutate(type = factor(type, levels = c("Cluster", "Clinic"))) -> clust_clinic

clust_clinic %>% 
  st_centroid() %>% 
  st_coordinates() %>% 
  as.data.frame() -> clust_cent

clust_clinic <- bind_cols(clust_clinic, clust_cent)

ggmap(blt_lines, 
      base_layer = ggplot(clust_clinic)) +
  geom_sf(aes(col = type, lty = type), fill = NA, lwd = 1) +
  geom_text(aes(label = clustid, x = X, y = Y)) +
  labs(x = "", y = "", col = "")
# ggsave(here::here("figures","cluster_clinic_map.png"), height = 6, width = 6)

```

## Missing data 

Primary analysis will be complete case, with multiple imputation as supplementary to check for robustness.

```{r missdata}

# Summarise missing data

dat %>% 
  dplyr::select(ever_test, s09age, sex, hiv_combined, pov_score, wealth_step) %>% 
  summary()

# Total incomplete observations
sum(is.na(dat$s09age) | dat$hiv_combined == "HIV unknown" | is.na(dat$pov_score))
# 148/2738 = 5.4%

dat_nona <- dat %>% 
  filter(!is.na(s09age) & hiv_combined != "HIV unknown" & !is.na(pov_score))

```

## Make analysis dataset

```{r make_data}

cluster_inc <- prev_clust %>% 
  st_drop_geometry() %>% 
  dplyr::select(cluster_number, clust_inc_1518)

dat_nona$hiv <- dat_nona$hiv_combined
dat_nona$hiv[dat_nona$hiv == "HIV positive ART"] <- "HIV positive"

dat_nona <- mutate(dat_nona, 
              test01 = as.numeric(ever_test == "Yes"),
              agegp = factor(cut(s09age, breaks = c(17,24,34,44,54,max(dat$s09age, na.rm=T)))),
              hiv = factor(hiv, levels = c("HIV negative","HIV positive"), labels = c("HIV negative","HIV positive")),
              pov_score_scale = -scale(pov_score),
              wealth_quant = as.factor(ntile(pov_score_scale, 6)),
              pov01 = as.factor(pov_score > 0),
              wealth_quant1 = as.factor(wealth_quant == 1),
              wealth_step = as.factor(wealth_step),
              wealth_step1 = as.factor(wealth_step == 1),
              know_test = as.factor(s60tb_anyone == 1),
              know_trt = as.factor(s62tb_famtb == 1),
              know_died = as.factor(s65tb_died == 1)) %>% 
  full_join(cluster_inc, by = c("clustid" = "cluster_number"))

# Check
dat_nona %>% 
  dplyr::select(test01, sex, agegp, hiv, pov_score, pov_score_scale, pov01, wealth_quant, wealth_step, wealth_step1, clust_inc_1518, clinic_id) %>% 
  summary()

```

# Modelling

## Priors

Weakly-informative

```{r priors}

priors <- c(prior(normal(-2, 2), class = Intercept),
            prior(normal(0,10), class = b),
            prior(cauchy(0,1), class = sd, group = "clustid"))

```

## Null model 

Fit a null model with no cluster structure or covariates. Weakly informative
prior on the intercept. 

```{r fit_simp}

fit0 <- brm(test01 ~ 1, 
                 data = dat_nona,
                 prior = prior(normal(-2, 2), class = "Intercept"),
                 family = "bernoulli")
summary(fit0)

# Prior predictive check
pp_check(fit0)

# Posterior predictive checks
y <- dat_nona$test01
yrep <- posterior_predict(fit0, draws = 20)

# Individual residuals
dat_nona %>%
  add_residual_draws(fit0, ndraws = 20) %>%
  ggplot(aes(x = .row, y = .residual)) +
  stat_pointinterval()

dat_nona %>%
  add_predicted_draws(fit0, ndraws = 20) %>%
  make_probability_residuals(.prediction, test01, n = 1) %>%
  ggplot(aes(sample = .p_residual)) +
  geom_qq(distribution = qunif) +
  geom_abline()

# Leave-one-out cross-validation
loo0 <- LOO(fit0, save_psis = TRUE)
plot(loo0)

```

Note on LOO for discrete outcomes:

https://discourse.mc-stan.org/t/worst-loo-pit-plots-ever-pp-checks-great/18789/3

LOO PIT expected to look wrong as not designed for discrete outcomes. 


## Fixed effect model

Note that in brm the intercept prior is applied assuming mean-centred covariates, i.e. the expected response when 
all covariates at their means. All predictors here are factors, so intercept is the expected response for a HIV- male aged 18-24. 

```{r fit_fixeff}

fit1 <- brm(test01 ~ 1 + agegp + sex + hiv, 
                 data = dat_nona,
                 prior = c(prior(normal(-2, 2), class = Intercept),
                           prior(normal(0,10), class = b)),
                 family = "bernoulli")
summary(fit1)

pairs(fit1)

# Prior predictive check
pp_check(fit1)

# Posterior predictive checks
y <- dat_nona$test01
yrep <- posterior_predict(fit1, draws = 20)

# Individual residuals
dat_nona %>%
  add_residual_draws(fit1, ndraws = 20) %>%
  ggplot(aes(x = .row, y = .residual)) +
  stat_pointinterval()

dat_nona %>%
  add_predicted_draws(fit1, ndraws = 20) %>%
  make_probability_residuals(.prediction, test01, n = 1) %>%
  ggplot(aes(sample = .p_residual)) +
  geom_qq(distribution = qunif) +
  geom_abline()

# Leave-one-out cross-validation
loo1 <- LOO(fit1, save_psis = TRUE)
plot(loo1)

```

```{r fit_fixeff_int}

fit1a <- brm(test01 ~ 1 + agegp*sex + hiv, 
                 data = dat_nona,
                 prior = c(prior(normal(-2, 2), class = Intercept),
                           prior(normal(0,10), class = b)),
                 family = "bernoulli")

summary(fit1a)

# Prior predictive check
pp_check(fit1a)

# Posterior predictive checks
y <- dat_nona$test01
yrep <- posterior_predict(fit1a, draws = 20)

# Individual residuals
dat_nona %>%
  add_residual_draws(fit1a, ndraws = 20) %>%
  ggplot(aes(x = .row, y = .residual)) +
  stat_pointinterval()

dat_nona %>%
  add_predicted_draws(fit1a, ndraws = 20) %>%
  make_probability_residuals(.prediction, test01, n = 1) %>%
  ggplot(aes(sample = .p_residual)) +
  geom_qq(distribution = qunif) +
  geom_abline()

# Leave-one-out cross-validation
loo1a <- LOO(fit1a, save_psis = TRUE)
plot(loo1a)

```

## Hierarchical model

Fit a regression model for ever tested, adjusted for age, sex and HIV status
with IID random effects by cluster. Compare with and without age*sex interaction.

```{r fit_clust}

# Note:
# If the posterior sd for any parameter (or qoi) is more than 0.1 times the prior sd, 
# then print out a note: "The prior distribution for this parameter is informative." 
# Then the user can go back and check that the default prior makes sense for this 
# particular example.

fit2 <- brm(test01 ~ sex + agegp + hiv + (1 | clustid), 
                 data = dat_nona,
                 prior = priors,
                 family = "bernoulli")

summary(fit2)

# Prior predictive check
pp_check(fit2)

pairs(fit2)

# Posterior predictive checks
y <- dat_nona$test01
yrep <- posterior_predict(fit2, draws = 20)

# Individual residuals
dat_nona %>%
  add_residual_draws(fit2, ndraws = 20) %>%
  ggplot(aes(x = .row, y = .residual)) +
  stat_pointinterval()

dat_nona %>%
  add_predicted_draws(fit2, ndraws = 20) %>%
  make_probability_residuals(.prediction, test01) %>%
  ggplot(aes(sample = .p_residual)) +
  geom_qq(distribution = qunif) +
  geom_abline()

# Leave-one-out cross-validation
loo1a <- LOO(fit2, save_psis = TRUE)
plot(loo1a)

# Conditional effects
# pdf(here::here(figdir, "condeff_base1.pdf"), height = 5, width = 6)  
plot(conditional_effects(fit2), ask = F)
# dev.off()

```

### With interaction

```{r fit_clust_int}

# With interaction
fit2a <- brm(test01 ~ sex*agegp + hiv + (1 | clustid), 
                 data = dat_nona,
                 prior = priors,
                 family = "bernoulli")
summary(fit2a)

# Prior predictive check
pp_check(fit2a)
pp_check(fit2a, type = "stat",stat = "mean", ndraws = 500)
# ggsave(here::here(figdir,"ppc_tstat_baseline.png"), height = 5, width = 5)

# Posterior predictive checks
y <- dat_nona$test01
yrep <- posterior_predict(fit2a, draws = 20)

# Individual residuals
dat_nona %>%
  add_residual_draws(fit2a, ndraws = 20) %>%
  ggplot(aes(x = .row, y = .residual)) +
  stat_pointinterval()

dat_nona %>%
  add_predicted_draws(fit2a, ndraws = 20) %>%
  make_probability_residuals(.prediction, test01) %>%
  ggplot(aes(sample = .p_residual)) +
  geom_qq(distribution = qunif) +
  geom_abline()

# Leave-one-out cross-validation
loo1a <- LOO(fit2a, save_psis = TRUE)
plot(loo2a)

# Conditional effects
# pdf(here::here(figdir, "condeff_base1.pdf"), height = 5, width = 6)  
plot(conditional_effects(fit2a), ask = F)
# dev.off()

```

### Plot baseline model coefficients

```{r plot_fit_base}

plot_model(fit2a, 
           order.terms = c(1,2,3,4,5,7,8,9,10,6),
           transform = NULL, 
           title = "",
           axis.labels = rev(c("Male","Age (24,34]", "Age (34,44]", "Age (44,54]", "Age (54,87]",
                           "Male : Age (24,34]","Male : Age (34,44]","Male : Age (44,54]", "Male : Age (54,87]",
                           "HIV positive"))
           ) + 
  geom_hline(yintercept = 0, lty = "dashed", col = "grey")
ggsave(here::here(figdir,"fixeff_baseline.png"), height = 6, width = 6)

```

## Save final baseline fit and analysis data

```{r save_base_fit}

saveRDS(dat_nona, here::here("data","analysis_data_primary.rds"))

saveRDS(fit2a, here::here("output","fit_base.rds"))

```