---
title: "Prev_survey_trial"
author: "Helena Feasey, Emily Nightingale"
date: "13/01/2022"
output: html_document
---

## 1. Backgound


<br>

## 2. Set-up

Load all required packages for analysis.

```{r setup, echo=FALSE, include=FALSE, message=FALSE, comment=NA, warning=FALSE}

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)    #for data manipulation
library(knitr)        #for tables
library(haven)
library(labelled)
library(lubridate)
library(sf)
library(janitor)
library(zoo)
library(epitools)
library(dplyr)
```

<br>

## 3. Import datasets

Import data required for the analysis.

```{r import, message=FALSE, warning=FALSE}

datadir <- "C:/Users/phpuenig/Dropbox/SCALE_Hit-TB_modelling-resources/ModellingData/Pre_PS_Data"

# individual_survey <- read_dta(file.path(datadir,"individual.dta"))
household_survey <- read_dta(file.path(datadir,"household_wide.dta")) %>% 
  haven::zap_formats() %>%
  haven::zap_label() %>%
  haven::zap_label() %>% 
  as_tibble()
screening <- read_dta(file.path(datadir, "ind_xray_hts.dta")) %>%
  haven::zap_formats() %>%
  haven::zap_label() %>%
  haven::zap_label() %>% 
  as_tibble()

```

<br>

## 4 Tidy

```{r tidy}
table(household_survey$arm, useNA = "always")

  
# Clean up household survey
household_survey <- household_survey %>%
  mutate(arm = case_when(arm==0 ~ "SOC",
                         arm==1 ~ "ACF"),
         arm = fct_relevel(arm, "ACF", "SOC"))


#Clean up demographic and clinical characteristics - individual + HTS survey - 'screening data'
screening <- screening %>%
mutate(sex = case_when(s07sex==1 ~ "Male",
                         s07sex==2 ~ "Female",
                         s07sex==99 ~ NA_character_),
         sex = fct_relevel(sex, "Female", "Male")) %>%
  mutate(arm = case_when(arm==0 ~ "SOC",
                         arm==1 ~ "ACF"),
         arm = fct_relevel(arm, "ACF", "SOC")) %>%
  mutate(TB_contact = case_when(s62tb_famtb==1 ~ "Yes",
                                 s62tb_famtb==2 ~ "No"),
         TB_contact = fct_relevel(TB_contact, "Yes", "No")) %>%
  mutate(previous_tb = case_when(s56tb_evrtst==1 ~ "Yes",
                                  s56tb_evrtst==2 ~ "No"),
         previous_tb = fct_relevel(previous_tb, "Yes", "No")) %>%
  mutate(hiv_ever = case_when(s72hiv_ever==1 ~ "Yes",
                                    s72hiv_ever==2 ~ "No",
                                    TRUE ~ NA_character_),
         hiv_ever = fct_relevel(hiv_ever, "Yes", "No")) %>%
  mutate(s74hiv_rectest = case_when(s74hiv_rectest==1 ~ "Positive",
                                    s74hiv_rectest==2 ~ "Negative",
                                    s74hiv_rectest==3 ~ "Inconclusive",
                                    s74hiv_rectest==4 ~ "Unknown",
                                    TRUE ~ NA_character_),
         s74hiv_rectest = fct_relevel(s74hiv_rectest, "Positive", "Negative", "Inconclusive", "Unknown")) %>%
  mutate(s80hiv_art = case_when(s80hiv_art==1 ~ "Yes",
                                    s80hiv_art==2 ~ "No",
                                    s80hiv_art==3 ~ "Can't Reveal",
                                    TRUE ~ NA_character_),
         s80hiv_art = fct_relevel(s80hiv_art, "Yes", "No", "Can't Reveal")) %>%
  mutate(s82hiv_crart = case_when(s82hiv_crart==1 ~ "Yes",
                                    s82hiv_crart==2 ~ "No",
                                    TRUE ~ NA_character_),
         s82hiv_crart = fct_relevel(s82hiv_crart, "Yes", "No")) %>%
  mutate(d12sumres = case_when(d12sumres==1 ~ "Previous positive not on ART",
                               d12sumres==2 ~ "Previous positive on ART",
                               d12sumres==3 ~ "New negative",
                               d12sumres==4 ~ "New positive",
                               d12sumres==5 ~ "Not done",
                               TRUE ~ NA_character_),
         d12sumres = fct_relevel(d12sumres, "Previous positive not on ART", "Previous positive on ART", "New negative", "New positive", "Not done"))
```
### Missing age

Where age is missing, infer from given DOB

```{r imp_age}

agemiss <- which(is.na(screening$s09age))
length(agemiss)

apply(dplyr::select(screening, s08dobday:s08dobyr), 2, summary)

screening[,c("s08dobday","s08dobmon","s08dobyr")] <- screening %>% 
  dplyr::select(s08dobday:s08dobyr) %>% 
  purrr::map2(c(88,77,888), function(x, y) na_if(x, y)) 

screening <- screening %>% 
  # If day of birth unknown but month/year known, impute with 1st of month
  # If month of birth unknown but year known, impute with June
  mutate(s08dobday = case_when((is.na(s08dobday) & !is.na(s08dobyr)) ~ 1,
                               T ~ s08dobday),
         s08dobmon = case_when((is.na(s08dobmon) & !is.na(s08dobyr)) ~ 6,
                               T ~ s08dobmon),
         dob_ = paste(s08dobday, s08dobmon, s08dobyr, sep = "."),
         dob = dmy(dob_),
         age=case_when(
           is.na(s09age) ~ floor(as.numeric(interval(start = dob, end = data_date) / 
                                              duration(num = 1, units = "years"))),
           T ~ s09age))

summary(screening$age)

# Define 10 year age bands
screening <- screening %>%
    mutate(agegp = case_when(between(age,18,24) ~ "18-24",
                           between(age,25,34) ~ "25-34",
                           between(age,35,44) ~ "35-44",
                           between(age,45,54) ~ "45-54",
                           between(age,55,64) ~ "55-64",
                           age > 64 ~ "65+" ),
          agegp = factor(agegp,levels = c("18-24","25-34","35-44",
                                          "45-54","55-64","65+")))

summary(screening$agegp)
```

## 5 Aggregating HIV information

Tidy HIV screening variables. 
Define overall HIV status according to screening result, self-reporting of previous testing and self-reporting of ART use. 

```{r tidy_hiv}

# Create overall HIV/ART status variable
table(screening$hiv_ever, useNA = "always")
table(screening$s74hiv_rectest, useNA = "always")
table(screening$s80hiv_art, useNA = "always")
table(screening$s82hiv_crart, useNA = "always")

table(screening$hiv_ever, screening$s80hiv_art, useNA = "always")
table(screening$s74hiv_rectest, screening$s80hiv_art, useNA = "always")
table(screening$s80hiv_art, screening$s82hiv_crart, useNA = "always")
table(screening$s74hiv_rectest, screening$s82hiv_crart, useNA = "always")

individual_survey <- screening

individual_survey <- individual_survey %>%
  mutate(hiv_art = case_when(s82hiv_crart=="Yes" ~ "HIV positive ART",
                             s82hiv_crart=="No" & s74hiv_rectest=="Positive"  ~ "HIV positive",
                             # s82hiv_crart=="No" & (s74hiv_rectest=="Negative" | s74hiv_rectest=="Positive")  ~ "HIV positive",
                             (s82hiv_crart =="No"| is.na(s82hiv_crart)) & s74hiv_rectest=="Negative" ~ "HIV negative",
         # is.na(s82hiv_crart) & s74hiv_rectest=="Negative" ~ "HIV negative",
         is.na(s82hiv_crart) & (s74hiv_rectest=="Inconclusive" | s74hiv_rectest=="Unknown") ~ "HIV unknown"),
         hiv_art = fct_relevel(hiv_art, "HIV positive ART", "HIV positive", "HIV negative", "HIV unknown"))

table(individual_survey$hiv_art, useNA = "always")

# Re-do d12sumres on basis of actual testing data
individual_survey <- individual_survey %>% mutate(rmb_hiv=case_when(
  d07hivst==1 & d11unig==1 ~ "old pos", # said HIV pos, unig pos
  d07hivst==1 & d11unig==2 ~ "neg*", # this is the one person who said HIV pos but unigold neg
  d07hivst!=1 & d09oraq==1 & d10detr==1 & d11unig==1 ~ "new pos", # new pos, all confirmed
  is.na(d07hivst) & d09oraq==1 & d10detr==1 & d11unig==1 ~ "new pos", # new pos, all confirmed. NA to HIV previous HIV status
  d07hivst!=1 & d09oraq==2 & d10detr==2 & is.na(d11unig) ~ "neg", # negative (oraq and determine, unigold not done)
  is.na(d07hivst) & d09oraq==2 & d10detr==2 & is.na(d11unig) ~ "neg", # negative (oraq and determine, unigold not done) - but didn't answer the HIV status question
  d05const==2 ~ "no test",
  d07hivst==2 & d09oraq==2 & d10detr==1 & d11unig==1 ~ "new pos*", # these are two people who were neg on oraquick, but pos on determine and unigold
  d07hivst==2 & d09oraq==1 & d10detr==2 & d11unig==2 ~ "neg*", # these are three people who were pos on oraquick, but neg on determine and unigold
  d07hivst==2 & d09oraq==1 & d10detr==1 & d11unig==2 ~ "indeterminate", # this is one person who was oraq and determine pos, but unigold neg
)) 

individual_survey %>% tabyl(rmb_hiv)
individual_survey %>% tabyl(rmb_hiv,d08art)

# Match d12sumres values
individual_survey <- individual_survey %>% mutate(hiv = case_when(
  (rmb_hiv=="old pos" & d08art==1) ~ "Previous positive on ART",
  (rmb_hiv=="old pos" & d08art==2) ~ "Previous positive not on ART",
    rmb_hiv=="new pos" ~ "New positive",
    rmb_hiv=="new pos*" ~ "New positive",
      rmb_hiv=="neg*" ~ "New negative",
      rmb_hiv=="neg" ~ "New negative",
      rmb_hiv=="indeterminate" ~ "New negative",
      rmb_hiv=="no test" ~ "Not done"),
          hiv = factor(hiv,levels = c("Previous positive on ART","Previous positive not on ART","New positive","New negative","Not done")))

individual_survey %>% tabyl(hiv)

# Create combined HIV variable based on testing data and if none using reported data
individual_survey <- individual_survey %>%
  mutate(hiv_combined = hiv) %>%
  mutate(hiv_combined = case_when(hiv=="Previous positive on ART" ~ "HIV positive ART",
                                  hiv=="Previous positive not on ART" ~ "HIV positive",
                                  hiv=="New negative" ~ "HIV negative",
                                  hiv=="New positive" ~ "HIV positive",
                              (hiv=="Not done" | is.na(hiv)) & hiv_art=="HIV positive ART" ~ "HIV positive ART",
                             (hiv=="Not done" | is.na(hiv)) & hiv_art=="HIV positive" ~ "HIV positive",
                             (hiv=="Not done" | is.na(hiv)) & hiv_art=="HIV negative" ~ "HIV negative",
                             (hiv=="Not done" | is.na(hiv)) & hiv_art=="HIV unknown" ~ "HIV unknown",
                             (hiv=="Not done" | is.na(hiv)) & is.na(hiv_art) ~ "HIV unknown"),
         hiv_combined = fct_relevel(hiv_combined, "HIV positive ART", "HIV positive", "HIV negative"))

table(individual_survey$hiv_combined, useNA = "always")

individual_survey %>% 
  dplyr::select(hiv_ever:s86hiv_attend, d12sumres, hiv_art:hiv_combined) %>% 
  dplyr::filter(hiv_combined == "HIV unknown") -> hiv_miss

summary(hiv_miss)

hiv_miss$hiv_combined2 = hiv_miss$hiv_combined

```

## 6 Tidy TB definition and symptoms

```{r tidy_tb}
# TB symptoms tidy
individual_survey <- individual_survey %>%
  mutate(night_sweats = case_when(s69tb_swts==1 ~ "Yes",
                                 s69tb_swts==2 ~ "No"),
         night_sweats = fct_relevel(night_sweats, "Yes", "No")) %>%
  mutate(weight_loss = case_when(s70tb_wght==1 ~ "Yes",
                                  s70tb_wght==2 ~ "No"),
         weight_loss = fct_relevel(weight_loss, "Yes", "No")) %>%
  mutate(fever = case_when(s71tb_fvr==1 ~ "Yes",
                                  s71tb_fvr==2 ~ "No"),
         fever = fct_relevel(fever, "Yes", "No")) %>%
  mutate(xray = case_when(x07xres==1 ~ "Normal",
                                  x07xres==2 ~ "Abnormal"),
         xray = fct_relevel(xray, "Abnormal", "Normal"))
 
         
# Chronic cough variable (2 weeks or more)
range(individual_survey$s68tb_cghdur[individual_survey$s68tb_cghdur_period == 4], na.rm=TRUE)

individual_survey <- individual_survey %>%
  mutate(chronic_cough = case_when(s68tb_cghdur_period==4 & s68tb_cghdur>0 ~ "Yes",
                                  s68tb_cghdur_period==3 & s68tb_cghdur>0 ~ "Yes",
                                  s68tb_cghdur_period==2 & s68tb_cghdur>0 ~ "Yes",
                                  s68tb_cghdur_period==1 & s68tb_cghdur>13 ~ "Yes",
                                  s68tb_cghdur_period==1 & s68tb_cghdur<14 ~ "No",
                                  is.na(s68tb_cghdur_period) ~ "No",
                                  is.na(s68tb_cghdur) ~ "No")) 


table(individual_survey$chronic_cough, useNA = "always")

# Any cough - individual survey or X-ray data
individual_survey %>% tabyl(s66tb_cough,x06cough)

individual_survey <- individual_survey %>%
  mutate(any_cough = case_when(s66tb_cough==1 ~ "Yes",
                                  x06cough==1 ~ "Yes",
                                  s66tb_cough==2 & x06cough==2 ~ "No",
                                  s66tb_cough==2 & is.na(x06cough) ~ "No"),
                               any_cough = fct_relevel(any_cough, "Yes", "No")) 

individual_survey %>% tabyl(any_cough,x06cough)
individual_survey %>% tabyl(any_cough,s66tb_cough)
individual_survey %>% tabyl(any_cough)
individual_survey %>% tabyl(any_cough,xray)

# Any TB symptom (cough any duration)
individual_survey <- individual_survey %>%
  mutate(any_tb_symp = if_else(any_cough=="Yes" | s69tb_swts==1 | s70tb_wght==1 | s71tb_fvr==1, "Yes","No"))
                                  
table(individual_survey$any_tb_symp, useNA = "always")

# Eligible for sputum submission (NB 2407 no X-ray. 98 of these had cough in survey & 2309 no cough)
individual_survey <- individual_survey %>%
  mutate(sputum_eligible = case_when(xray=="Abnormal" ~ "Yes",
                                     any_cough=="Yes" ~ "Yes",
                                     xray=="Normal" & any_cough=="No" ~ "No",
                                     any_cough=="No" & is.na(xray) ~ "No"))

individual_survey %>% tabyl(sputum_eligible)
# individual_survey %>% tabyl(any_cough)
# individual_survey %>% tabyl(x06cough)
# individual_survey %>% tabyl(any_cough,s66tb_cough)
# individual_survey %>% tabyl(x06cough,s66tb_cough)
# individual_survey %>% tabyl(any_cough,any_tb_symp)

```

<br>

## 7 Tidy testing history

```{r last_test}

dat_na <- function(x){
  x[grepl("NA",x)] <- NA
  return(x)
}


individual_survey %>%
  dplyr::select(s52tb_stsptmon, s52tb_lstsptyr, s55tb_lstsxraymon, s55tb_lstxraytyr) %>% 
  summary()

individual_survey <- individual_survey %>%  
  mutate(# Ever test
         ever_test = factor((s50tb_medtest == 1 | s53tb_evrxray == 1), 
                               levels = c(FALSE,TRUE), 
                               labels = c("No","Yes")),
         # Last test
         last_sputum_m = s52tb_stsptmon,
         last_sputum_y = s52tb_lstsptyr,
         last_xray_m = s52tb_stsptmon,
         last_xray_y = s52tb_lstsptyr,
         # Define month as June if year is known
         last_sputum_m = replace(s52tb_stsptmon, 
                                  (s52tb_stsptmon == 77 & !s52tb_lstsptyr %in% c(888,999)), 
                                  6),
         last_xray_m = replace(s55tb_lstsxraymon, 
                                     (s55tb_lstsxraymon == 77 & !s55tb_lstxraytyr %in% c(888,999)), 
                                     6),
         # Otherwise, redefine missing codes as NA
         across(c(s52tb_stsptmon, s55tb_lstsxraymon), 
                function(x) na_if(x, 77)),
         across(c(s52tb_lstsptyr, s55tb_lstxraytyr), 
                function(x) na_if(na_if(x, 888),999)), 
         # Paste together (imputed) day/month/year and convert to date
         last_sputum = paste(1, s52tb_stsptmon, s52tb_lstsptyr, sep = "/"),
         last_xray = paste(1, s55tb_lstsxraymon, s55tb_lstxraytyr, sep = "/"),
         across(c(last_sputum, last_xray), dat_na),
         across(c(last_sputum, last_xray), dmy),
         # Identify last test and whether it occurred within 12m prior to interview. 
         # If last date = NA define outcome as false
         last_test = pmax(last_sputum, last_xray, na.rm = T),
         last_test_12m = factor(replace_na(
           (interval(last_test, today) %/% months(1) < 12), 
           FALSE))) 

individual_survey %>% 
  filter(ever_test == "Yes") %>% 
  pull(last_test) %>% 
  summary()
#         Min.      1st Qu.       Median         Mean      3rd Qu.         Max.         NA's 
# "1995-05-01" "2018-01-01" "2019-03-01" "2017-07-18" "2019-09-01" "2019-12-01"        "260" 

```

## 5 Overcrowding

```{r table}
table(household_survey$h28rm_roomnum)
table(household_survey$h07hh_number)
table(household_survey$h08hh_adults_num)

# Mean hh_per_room by arm
household_survey <- household_survey %>%
  mutate(hh_per_room = h07hh_number / h28rm_roomnum)


household_survey_arm <- household_survey %>%
  dplyr::group_by(arm) %>%
  dplyr::summarize(Mean = mean(hh_per_room))

household_survey_arm %>%
  tabyl(arm, Mean)

t.test(hh_per_room ~ arm, data=household_survey)

```
<br>

## 6 Poverty score

```{r poverty}

table(household_survey$h13hh_level_qual)
table(household_survey$h14hh_sleep)

#Create the poverty score
#First make a look-up table with variables and coefficients 
pov_lookup <- tribble(~var,            ~value,   ~coef,
                       "h13hh_level_qual",  1,   0,
                       "h13hh_level_qual",  2,   -0.1070230,
                       "h13hh_level_qual",  3, -0.4727860,
                       "h13hh_level_qual", 4, -0.3210555,
                       "h13hh_level_qual", 5, -14.1811944,
                       
                       "h14hh_sleep", 1, 0,
                       "h14hh_sleep", 2, 0.6709013,
                       "h14hh_sleep", 3, 0.3235319,
                       "h14hh_sleep", 4, 0.4599334,
                       "h14hh_sleep", 5, 0.7354924,
                       "h14hh_sleep", 6, 2.9858984,
                       "h14hh_sleep", 7, 0,
                       "h14hh_sleep", 77, -0.6208952,
                       
                       "h15hh_electric", 1, 0, 
                       "h15hh_electric", 2, 0.9590447,
                       
                       "h16hh_bank", 1, 0,
                       "h16hh_bank", 2, 0.8831791,
                       
                       "h17hh_food", 1, 0,
                       "h17hh_food", 2, -0.5573320,
                      
                       "h18hh_clothing", 1, 0,
                       "h18hh_clothing", 2, -0.6375926,
                       "h18hh_clothing", 3, -0.9265472,
                       
                       "h19hh_hifi", 1, 0,
                       "h19hh_hifi", 2, 0.8729641,
                       
                       "h20hh_sofa", 1, 0,
                       "h20hh_sofa", 2, 0.9942764,
                       
                       "h21hh_iron", 1, 0,
                       "h21hh_iron", 2, 1.3393567
                       )

#Now calculate poverty scores and merge to baseline data
temp_base <- household_survey %>%
  dplyr::select(hh_id, h13hh_level_qual:h21hh_iron) %>%
  dplyr::select(-h14hh_sleep_oth) %>%
  gather(var, value, -hh_id)

temp_base_b <- household_survey %>%
  dplyr::select(hh_id, h07hh_number) %>%
  mutate(h07hh_number = as.integer(h07hh_number)) %>%
  mutate(hhsize_coef = 0.6222069 * h07hh_number) %>%
  dplyr::select(-h07hh_number)

temp_base <- left_join(temp_base, pov_lookup)

temp_base <- temp_base %>%
  dplyr::select(-value) %>%
  spread(var, coef)

temp_base <- left_join(temp_base, temp_base_b)

temp_base <- temp_base %>%
  mutate(pov_score = -8.5761747 +  (rowSums(select_(., "-hh_id")))) %>%
  dplyr::select(hh_id, pov_score)

household_survey <- left_join(household_survey, temp_base)

rm(temp_base, temp_base_b)

# Poverty score by arm
household_poverty_arm <- household_survey %>%
  dplyr::group_by(arm) %>%
  dplyr::summarize(Mean = mean(pov_score))

t.test(pov_score ~ arm, household_survey)

# Generate wealth quartiles
household_survey <- household_survey %>%
          mutate(
            wealth_quantil = ntile(pov_score, 4) 
          )

table(household_survey$arm, household_survey$wealth_quantil)
wealth_quartileT<-table(household_survey$arm, household_survey$wealth_quantil)
round(prop.table(wealth_quartileT,margin=1)*100,digits=1)
chisq.test(wealth_quartileT)

```

<br>

```{r link_clinic}

# Link clusters to clinics

clust_clinic <- read_csv(here::here("data","PreRandomisationClustData_23May2016.csv")) %>% 
  mutate(s02cl_id = as.numeric(gsub("C","",clid)),
         clinic_id = as.numeric(as.factor(close_clin))) %>% 
  dplyr::select(clinic_id, close_clin, s02cl_id)

individual_survey <- individual_survey %>% 
  full_join(clust_clinic) 

```

```{r save_datasets}

individual_survey <- ungroup(individual_survey)
household_survey <- ungroup(household_survey)

survey <- individual_survey %>% 
  left_join(dplyr::select(household_survey, 
                          hh_id, h02cl_id, h44gpsid:gps_acc, 
                          total_hhmember:wealth_quantil, h22hh_step))

saveRDS(individual_survey, file.path(datadir, "individual_clean.rds"))
saveRDS(household_survey, file.path(datadir, "household_clean.rds"))
saveRDS(survey, file.path(datadir, "survey_clean.rds"))

```

```{r setup_analysis_data}

# Individual data
dat <- survey %>% 
  mutate(any_tb_symp = as.factor(any_tb_symp),
         current_trt = factor((s57tb_treat == 1), 
                              levels = c(FALSE,TRUE), 
                              labels = c("No","Yes")),
         last_test_12m = factor(last_test_12m, levels = c(FALSE,TRUE), labels = c("No","Yes")),
         incl_extended = !is.na(s50tb_medtest)) %>% 
  rename(clustid = s02cl_id,
         wealth_step = h22hh_step) %>% 
  # Count participants per cluster
  group_by(clustid) %>% 
  mutate(clust_N_tot = n(),
         clust_N_ext = sum(incl_extended)) %>% 
  ungroup() 

saveRDS(dat, file.path(datadir, "analysis_data.rds"))

```
